---
title: "Comparisons simulated interaction effect size for 3D sim"
date: today
format:
  html:
    embed-resources: true
    code-fold: true
    toc: true
    toc-location: left
execute: 
  cache: false
  warning: false
  error: false
  fig-format: png
knitr:
  opts_chunk:
    dev: png
---




```{r}
library(tidyverse)
library(patchwork)
library(ggsci)
library(gt)
dir.create("tables", showWarnings = FALSE, recursive = TRUE)
# helper: wrap gt in scrollable div (horizontal + vertical)
gt_scroll <- function(gt_tbl, max_h = "500px") {
  htmltools::div(style = sprintf("overflow-x:auto; overflow-y:auto; max-height:%s; border:1px solid #ddd; padding:4px;", max_h), gt_tbl)
}
theme_set(theme_bw())
theme_update(
    scale_fill = scale_fill_npg(),
    scale_color = scale_color_npg(),
)

df <- data.frame("file" = dir("../results2")) %>%
    filter(grepl(".csv$", file), !grepl("^covariates", file)) %>%
    mutate(
        Imbalanced = grepl("^Imbal_", file),
        mass1 = str_extract(file, "mass[0-9].0"),
        mass1 = as.numeric(str_remove(mass1, "^mass")),
        mass2 = str_extract(file, "[0-9].0.csv"),
        mass2 = as.numeric(str_remove(mass1, ".csv$")),
    ) %>%
    # map read these csvs into nested dataframe
    mutate(data = map(file, ~ read_csv(file.path("../results2", .x))))  %>%
    # unnest them
    unnest(data) %>%
    select(
        `ppmx-common_rind` = adjrind_Mix,
        ppmx_rind = adjrind_DPM,
        dpm_rind = dpmARI,
        kmean_rind = adjrind_K,
        `ppmx-common_arind` = adjrindMixoos,
        ppmx_arind = adjrind_DPMoos,
        dpm_arind = dpmARIoos,
        kmean_arind = adjrind_Koos,

        `ppmx-common_rmse` = midMixoos,
        ppmx_rmse = midDPMoos,
        dpm_rmse = dpmRMSEoos,
        kmean_rmse = kmean_MSEoos,
        slr_rmse = slrRMSEoos,
        
        `ppmx-common_common1` = meanBeta1,
        `ppmx-common_common2` = meanBeta2,
        dpm_common1 = dpmBetaclust1,
        dpm_common2 = dpmBetaclust2,
        kmean_common1 = meanBetakclust1,
        kmean_common2 = meanBetakclust2,
        slr_common1 = meanBetaSLR,
        slr_common2 = meanBetaSLR2,

        `ppmx-common1_T2` = zeroInDPM,
        `ppmx-common2_T2` = zeroInDPM2,
        dpm1_T2 = dpmzeroIn1,
        dpm2_T2 = dpmzeroIn2,
        kmean1_T2 = zeroInk,
        kmean2_T2 = zeroInk2,
        slr1_T2 = zeroInSLR,
        slr2_T2 = zeroInSLR2,
        `ppmx-common1_Cov` = commonInDPM,
        `ppmx-common2_Cov` = commonInDPM2,
        `ppmx-common_ccov` = commonCovAll,
        dpm1_Cov = dpmcommonIn1,
        dpm2_Cov = dpmcommonIn2,
        kmean1_Cov = commonInk,
        kmean2_Cov = commonInk2,
        slr1_Cov = commonInSLR,
        slr2_Cov = commonInSLR2,

        `ppmx-common_NC` = ncMix,
        `ppmx_NC` = ncDPM,
        dpm_NC = dpmnclusts,
        kmean_NC = ncK,
        
        `ppmx-common_bayesP` = bayesPmix,
        `ppmx_bayesP` = bayesPDPM,
        `ppmx-common_lps` = lpsMixoos,
        `ppmx_lps` = lpsDPMoos,

        `ppmx-common_binder` = salsoBinderARI_Mix,
        `ppmx-common_vi` = salsoVIARI_Mix,
        ppmx_binder = salsoBinderARI_DPM,
        ppmx_vi = salsoVIARI_DPM,
        `ppmx-common_binderoos` = salsoBinderARI_Mixoos,
        `ppmx-common_vioos` = salsoVIARI_Mixoos,
        ppmx_binderoos = salsoBinderARI_DPMoos,
        ppmx_vioos = salsoVIARI_DPMoos,

        # DPMM.jl baseline (`dpm`) has a single hard partition, so it has no
        # MCMC SALSO point estimate; surface its single-partition ARI vs truth
        # in the SALSO section for comparison.
        dpm_binder = dpmARI,
        dpm_vi = dpmARI,
        dpm_binderoos = dpmARIoos,
        dpm_vioos = dpmARIoos,

        contains("rind"),
        contains("bayes"),

        Common = common, Precision = prec, Alpha = alph,
        Beta = bet, InterEffect = interEffect, Imbalanced,
        xdiff, variance, N, mass1, mass2, dims
    )
# group by file
modLoc <- df %>%
    # select(-contains("rind"), -contains("bayes")) %>%
    reframe(
        # type 2 error
        across(ends_with("T2"), function(x) mean(x, na.rm = T)),
        # coverage
        across(ends_with("_Cov"), function(x) mean(x, na.rm = T)),
        # bias
        across(ends_with("_common1"), function(x) mean(Common - x, na.rm = T), .names= "{.col}-bias"),
        across(ends_with("_common2"), function(x) mean(-Common - x, na.rm = T), .names= "{.col}-bias"),
        # MSE
        across(ends_with("_common1"), function(x) sqrt(mean((Common - x)^2, na.rm = T)), .names= "{.col}-RMSE"),
        across(ends_with("_common2"), function(x) sqrt(mean((-Common - x)^2, na.rm = T)), .names= "{.col}-RMSE"),
        .by = c(N, Common, InterEffect, Alpha, Beta, Precision, Imbalanced, variance, xdiff, dims)
    ) %>%
    pivot_longer(
        -c(Common, Alpha, Beta, Precision, InterEffect, Imbalanced, variance, xdiff, dims),
         names_to = c(".value", "Metric"),
         names_pattern = "(.*)_(.*)",
        ) %>%
    select(-c(slr1, slr2, kmean1, kmean2, dpm1, dpm2, `ppmx-common1`, `ppmx-common2`)) %>%
    # filter out global assesses
    drop_na(`ppmx-common`, kmean, slr)


modLoc |>
  filter(xdiff %in% c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0), dims == 5) |>
  reframe(.by = c(xdiff, Metric), across(c(`ppmx-common`, dpm, kmean, slr), mean))


    
modGlob <- df %>%
    # select(-contains("rind"), -contains("bayes")) %>%
    reframe(
        across(ends_with("_bayesP"), function(x) mean(x, na.rm = T), .names= "{.col}-bayesP"),
        across(ends_with("_rind"), function(x) mean(x, na.rm = T), .names= "{.col}-rind"),
        .by = c(N, Common, InterEffect, Alpha, Beta, Precision, Imbalanced, variance, xdiff, dims)
    ) %>%
    pivot_longer(
        -c(Common, Alpha, Beta, Precision, InterEffect, Imbalanced, variance, xdiff, dims),
         names_to = c(".value", "Metric"),
         names_pattern = "(.*)_(.*)",
        ) %>%
    drop_na()
longDF <- df %>%
    pivot_longer(
        -c(N, Common, Alpha, Beta, Precision, InterEffect, Imbalanced, variance, xdiff, dims),
         names_to = c(".value", "Metric"),
         names_pattern = "(.*)\\_(.*)",
        ) %>%
    pivot_longer(c(`ppmx-common`, ppmx, dpm, kmean, slr), names_to = "Model", values_to = "Statistic") # %>%
    # filter(Metric != "rind" | Statistic >=0.4)


df |>
  filter(xdiff %in% c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0), dims == 5) |>
  reframe(.by = xdiff, across(ends_with("_Cov"), mean))

modGlob |>
  filter(xdiff %in% c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0), dims == 5)

cat("\n=== DATA AVAILABILITY for new sections (SALSO / LPS) ===\n")
longDF %>%
  filter(Metric %in% c("binder", "vi", "binderoos", "vioos", "lps")) %>%
  reframe(.by = c(Model, Metric), n_nonmiss = sum(!is.na(Statistic)), n_total = n()) %>%
  print()

```

# local interp
```{r}
#| column: page
locTbl <- modLoc |> 
    arrange(InterEffect, Alpha, Beta, Metric) |>
    mutate(
        group_label = sprintf("InterEffect = %s, Alpha = %s, Beta = %s, xdiff = %s", InterEffect, Alpha,Beta, xdiff ),
        parameter= ifelse(grepl("1", Metric), "1", "2"),
        Metric = ifelse(grepl("bias", Metric), "bias", "RMSE"),
    ) |>
    group_by(group_label) |>
    select(-c(InterEffect, Alpha, Beta, Precision)) |>
    filter(xdiff %in% c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0), dims == 5)
locTbl |> write_csv("tables/table_local_interp.csv")
locTbl |>
  gt() |>
  tab_header(title = "Local interpretation: bias / RMSE by xdiff") |>
  opt_interactive(use_compact_mode = TRUE) |>
  gt_scroll()
```


# Common Effect
```{r}
# longDF %>%
#     filter(Metric == "common", Imbalanced == TRUE) %>%
#     ggplot(aes(x = Statistic, fill = Model)) +
#     geom_density(alpha = 0.3) +
#     geom_vline(xintercept = 1.0) +
#     facet_grid(
#         rows = vars(InterEffect), cols = vars(Common, variance), scales = "free",
#         labeller = labeller(.rows = label_both, .cols = label_both)
#     ) +
#     ggtitle("Common effects sampling distribution (alpha=2, beta=1,prec=0.01), Imbalanced") +
#     xlab("Common Estimate")
longDF %>%
    filter(Metric == "common1", dims == 5, Common == 1.0) %>%
    ggplot(aes(x = Statistic, fill = Model)) +
    geom_density(alpha = 0.35) +
    geom_vline(aes(xintercept = unique(Common))) +
    geom_vline(xintercept = 0.0) +
    facet_grid(
        rows = vars(xdiff, Imbalanced), cols = vars(Alpha, Beta), scales = "free",
        labeller = labeller(.rows = label_both, .cols = label_both)
    ) +
    ggtitle("Common effects (common=1, Interaction=1), dims=5") +
    xlab("Common Estimate") +
    xlim(-3, 3)
ggsave("beta1Tuning.png", width = 6, dpi = 300, units = "in")

longDF %>%
    filter(Metric == "common2", dims == 5, Common == 1.0) %>%
    ggplot(aes(x = Statistic, fill = Model)) +
    geom_density(alpha = 0.35) +
    geom_vline(aes(xintercept = -unique(Common))) +
    geom_vline(xintercept = 0.0) +
    facet_grid(
        rows = vars(xdiff, Imbalanced), cols = vars(Alpha, Beta), scales = "free",
        labeller = labeller(.rows = label_both, .cols = label_both)
    ) +
    ggtitle("Common effects (common=-1, Interaction=1), dims=5") +
    xlab("Common Estimate") +
    xlim(-3, 3)
ggsave("beta2Tuning.png", width = 6, dpi = 300, units = "in")
```

```{r}
longDF %>%
    filter(grepl("^common", Metric)) %>%
    ggplot(aes(x = Statistic, fill = Model)) +
    geom_density(alpha = 0.35) +
    geom_vline(xintercept = 0.0) +
    # geom_vline(aes(xintercept = unique(Common))) +
    facet_grid(
        rows = vars(xdiff, Alpha, Beta, Imbalanced), cols = vars(Metric), scales = "free",
        labeller = labeller(.rows = label_both, .cols = label_both)
    ) +
    ggtitle("Common effects (common=5, Interaction=5)") +
    xlab("Common Estimate") +
    xlim(-3, 3)
ggsave("betaVinter.png", width = 6, dpi = 300, units = "in")
```


## Model comparison tables 
- tuning InterEffect = 0.25, variance = 0.1, Common = 1.0
```{r}
#| column: page
modCompTbl <- longDF %>%
    filter(dims == 5, xdiff %in% c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0)) %>%
    group_by(Metric, Model, Imbalanced) %>%
    summarise(Statistic = mean(Statistic, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = Model, values_from = Statistic)
modCompTbl |> write_csv("tables/table_model_comparison.csv")
modCompTbl |>
  gt() |>
  tab_header(title = "Model comparison (dims=5, xdiff grid)") |>
  fmt_number(decimals = 3, use_seps = FALSE) |>
  opt_interactive(use_compact_mode = TRUE, use_page_size_select = TRUE) |>
  gt_scroll()
```

```{r}
#| column: page
tempMetricsTbl <- longDF %>%
    filter(dims == 5) %>%
    group_by(Model, Metric) %>%
    summarise(Statistic = mean(Statistic, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = Metric, values_from = Statistic)
tempMetricsTbl |> write_csv("tables/table_temp_metrics.csv")
tempMetricsTbl |>
  gt() |>
  tab_header(title = "Metrics by model (dims=5)") |>
  fmt_number(decimals = 3, use_seps = FALSE) |>
  opt_interactive(use_compact_mode = TRUE, use_page_size_select = TRUE) |>
  gt_scroll()
```


- vs InterEffects
```{r}
#| column: page
vsInterTbl <- longDF %>%
    filter(dims == 5, Alpha == 3.0, Beta == 3.0) %>%
    group_by(Model, Metric, InterEffect) %>%
    summarise(Statistic = mean(Statistic, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = Metric, values_from = Statistic)
vsInterTbl |> write_csv("tables/table_temp_vs_intereffects.csv")
vsInterTbl |>
  gt() |>
  tab_header(title = "Metrics vs InterEffect (Alpha=3, Beta=3, dims=5)") |>
  fmt_number(decimals = 3, use_seps = FALSE) |>
  opt_interactive(use_compact_mode = TRUE, use_page_size_select = TRUE) |>
  gt_scroll()
```

## clustering and prediction vs inter effect
```{r}
longDF %>%
    filter(dims == 5, Alpha == 3.0, Beta == 3.0, grepl("rind", Metric) | grepl("bayes", Metric)) %>%
    select(-c(Alpha, Beta, Precision, variance, Common, Imbalanced)) %>%
    arrange(InterEffect, Metric)
```


```{r}
longDF %>%
    filter(dims == 5, Metric == "rind") %>%
    ggplot(aes(x = Statistic, fill = Model, y = after_stat(density))) +
    geom_histogram(alpha = 0.3, color = "black") +
    facet_grid(cols = vars(Alpha), rows = vars(Beta), scales = "free",
        labeller = labeller(.rows = label_both, .cols = label_both)
    ) +
    xlim(0.8, 1.0) +
    ggtitle("Rand index (common=1, Interaction=0.25), Balanced")
```

# covariate distributions
```{r}
if (file.exists("covariates.csv")) {
covs <- read_csv("covariates.csv", show_col_types = FALSE) %>%
    pivot_longer(starts_with("X"), names_to = "covariate", values_to = "value")

covs %>%
    ggplot(aes(x = value, fill = factor(group))) +
    geom_density(alpha = 0.35, linewidth = 0.4) +
    facet_grid(
        rows = vars(covariate), cols = vars(xdiff),
        labeller = labeller(xdiff = function(l) sprintf("xdiff = %s", l))
    ) +
    scale_fill_npg() +
    labs(
        x = "Covariate value",
        y = "Density",
        fill = "Group"
    ) +
    theme_minimal(base_size = 11) +
    theme(
        legend.position = "right",
        strip.text = element_text(size = 9),
        axis.title.x = element_text(margin = margin(t = 8))
    ) +
    guides(fill = guide_legend(ncol = 1))
ggsave("covariateDist.png", width = 10, height = 8, dpi = 300, units = "in")
} else {
  cat("covariates.csv not found; run: julia --project=. covDist.jl\n")
}
```

# big sim
```{r}
longDF %>%
    filter(Metric == "common1", N== 1000, dims == 5, Common == 1.0) %>%
    ggplot(aes(x = Statistic, fill = Model)) +
    geom_density(alpha = 0.2) +
    geom_vline(xintercept = 1.0) +
    geom_vline(xintercept = 0.0) +
    facet_grid(
        rows = vars(xdiff, Imbalanced), cols = vars(InterEffect), scales = "free",
        #labeller = labeller(.rows = label_both, .cols = label_both)
    ) +
    ggtitle("Common effects, dims=5") +
    xlab("Common Estimate") +
    xlim(-3, 3)
ggsave("bigSimcommon1.png", width = 6, dpi = 300, units = "in")

longDF %>%
    filter(Metric == "common2", N== 1000, dims == 5, Common == 1.0) %>%
    ggplot(aes(x = Statistic, fill = Model)) +
    geom_density(alpha = 0.2) +
    geom_vline(xintercept = 1.0) +
    geom_vline(xintercept = 0.0) +
    facet_grid(
        rows = vars(xdiff, Imbalanced), cols = vars(InterEffect), scales = "free",
        #labeller = labeller(.rows = label_both, .cols = label_both)
    ) +
    ggtitle("Common effects, dims=5") +
    xlab("Common Estimate") +
    xlim(-3, 3)
ggsave("bigSimcommon2.png", width = 6, dpi = 300, units = "in")

longDF %>%
    filter(Metric == "rind", N == 1000, dims == 5, Common == 1.0) %>%
    ggplot(aes(x = Statistic, fill = Model, y = after_stat(density))) +
    geom_histogram(alpha = 0.3, color = "black") +
    facet_grid(rows = vars(xdiff), cols = vars(Imbalanced)) + #, labeller = labeller(.rows = label_both, .cols = label_both))
    ggtitle("Rand index, N = 1000, dims=5")
ggsave("bigSimRI.png", width = 6, dpi = 300, units = "in")

longDF %>%
    filter(Metric %in% c("binder", "vi", "binderoos", "vioos"), N == 1000, dims == 5, Common == 1.0) %>%
    mutate(Split = ifelse(grepl("oos", Metric), "OOS", "in-sample"),
           Loss = ifelse(grepl("binder", Metric), "Binder", "VI")) %>%
    ggplot(aes(x = Statistic, fill = Model, y = after_stat(density))) +
    geom_histogram(alpha = 0.3, color = "black") +
    facet_grid(rows = vars(Loss, Split), cols = vars(xdiff, Imbalanced)) +
    labs(x = "Adjusted Rand Index (SALSO point estimate)") +
    ggtitle("SALSO ARI, N = 1000, dims=5")
ggsave("bigSimSalso.png", width = 8, dpi = 300, units = "in")

```

# SALSO summary table (Binder's loss + VI error, in-sample and OOS)
The PPMx models (`ppmx-common`, `ppmx`/DPM) report SALSO point estimates under Binder
and VI loss. The DPMM.jl baseline (`dpm`) yields a single hard partition, so it
contributes its single-partition ARI versus truth (shown under Binder/VI for
comparison, not a SALSO point estimate).
```{r}
#| column: page
salsoTbl <- longDF %>%
    filter(Metric %in% c("binder", "vi", "binderoos", "vioos"), N == 1000, dims == 5, Common == 1.0) %>%
    mutate(Split = ifelse(grepl("oos", Metric), "OOS", "in-sample"),
           Loss  = ifelse(grepl("binder", Metric), "Binder", "VI")) %>%
    reframe(.by = c(Model, Loss, Split, xdiff, Imbalanced),
            ARI = mean(Statistic, na.rm = TRUE)) %>%
    pivot_wider(names_from = c(Loss, Split), values_from = ARI)
salsoTbl |> write_csv("tables/table_salso_by_model.csv")
salsoTbl |>
  gt() |>
  tab_header(title = "SALSO ARI by model (Binder / VI, in-sample & OOS)") |>
  fmt_number(decimals = 3, use_seps = FALSE) |>
  opt_interactive(use_compact_mode = TRUE, use_page_size_select = TRUE) |>
  gt_scroll()
```

# OOS log predictive score (reviewer: replace Bayesian p-values with proper scoring rules)
```{r}
#| column: page
lpsTbl <- longDF %>%
    filter(Metric == "lps", N == 1000, dims == 5, Common == 1.0) %>%
    reframe(.by = c(xdiff, Model, Imbalanced), Statistic = mean(Statistic, na.rm = TRUE)) %>%
    pivot_wider(names_from = Model, values_from = Statistic) %>%
    mutate(across(c(`ppmx-common`, ppmx), round, 3))
lpsTbl |> write_csv("tables/table_lps_by_model.csv")
lpsTbl |>
  gt() |>
  tab_header(title = "OOS log predictive score by model (N=1000, dims=5)") |>
  fmt_number(decimals = 3, use_seps = FALSE) |>
  opt_interactive(use_compact_mode = TRUE, use_page_size_select = TRUE) |>
  gt_scroll()
```

```{r}
longDF %>%
    filter(Metric == "lps", N == 1000, dims == 5, Common == 1.0) %>%
    ggplot(aes(x = Statistic, fill = Model, y = after_stat(density))) +
    geom_histogram(alpha = 0.3, color = "black") +
    facet_grid(rows = vars(xdiff), cols = vars(Imbalanced)) +
    ggtitle("OOS log predictive score, N = 1000, dims=5")
ggsave("bigSimLPS.png", width = 6, dpi = 300, units = "in")

# direct model comparison of OOS log predictive score (LPS)
longDF %>%
    filter(Metric == "lps", N == 1000, dims == 5, Common == 1.0) %>%
    ggplot(aes(x = Model, y = Statistic, fill = Model)) +
    geom_boxplot(alpha = 0.6) +
    facet_grid(rows = vars(xdiff), cols = vars(Imbalanced)) +
    labs(x = "Model", y = "OOS log predictive score (LPS)",
         title = "Out-of-sample log predictive score by model, N = 1000, dims = 5") +
    theme_bw()
ggsave("bigSimLPScompare.png", width = 8, dpi = 300, units = "in")
```

## Sensitivity to the Dirichlet process concentration hyperparameter (Rev 3 C2)
The simulation varies the DP concentration `massParams = [A, B]`, which controls the
prior expected number of clusters. The main `dims = 5` comparisons above were run at a
single mass value (`A = 1.0`); the table and figure below summarise the key
fit/partition metrics across the values of `A` (labelled `mass1`) available in the
`dims = 5` data, averaged over simulation replicates, to show the conclusions are
stable to this hyperparameter.

```{r}
#| column: page
dp2 <- df %>% filter(dims == 5, !is.na(mass1))
if (nrow(dp2) > 0) {
  dpSens <- dp2 %>%
      select(mass1, Imbalanced, xdiff,
             `ppmx-common_rind`, ppmx_rind, dpm_rind, kmean_rind,
             `ppmx-common_rmse`, ppmx_rmse, dpm_rmse, kmean_rmse, slr_rmse,
             `ppmx-common_NC`, ppmx_NC, dpm_NC, kmean_NC,
             `ppmx-common_binder`, ppmx_binder, `ppmx-common_vi`, ppmx_vi,
             `ppmx-common_lps`, ppmx_lps) %>%
      pivot_longer(-c(mass1, Imbalanced, xdiff),
                   names_to = c("Model", "Metric"),
                   values_to = "Statistic",
                   names_pattern = "(.*)_(.*)") %>%
      group_by(mass1, Model, Metric) %>%
      summarise(Statistic = mean(Statistic, na.rm = TRUE), .groups = "drop") %>%
      pivot_wider(names_from = Metric, values_from = Statistic)
  dpSens |> write_csv("tables/table_dp_hyperparameter_sensitivity.csv")
  dpSens |>
    gt() |>
    tab_header(title = "Sensitivity to DP concentration (mass A), dims=5") |>
    fmt_number(decimals = 3, use_seps = FALSE) |>
    opt_interactive(use_compact_mode = TRUE, use_page_size_select = TRUE) |>
    gt_scroll() |> print()

  dpSensFig <- dp2 %>%
      select(mass1,
             `ppmx-common_rind`, ppmx_rind, dpm_rind, kmean_rind,
             `ppmx-common_rmse`, ppmx_rmse, dpm_rmse, kmean_rmse, slr_rmse,
             `ppmx-common_NC`, ppmx_NC, dpm_NC, kmean_NC) %>%
      pivot_longer(-mass1, names_to = c("Model", "Metric"),
                   values_to = "Statistic",
                   names_pattern = "(.*)_(.*)") %>%
      filter(Metric %in% c("rind", "rmse", "NC")) %>%
      mutate(Metric = recode(Metric,
                             rind = "Adjusted Rand Index",
                             rmse = "OOS RMSE",
                             NC   = "Number of clusters")) %>%
      ggplot(aes(x = mass1, y = Statistic, color = Model, group = Model)) +
      geom_point(size = 2) + geom_line() +
      facet_wrap(~ Metric, scales = "free_y") +
      labs(x = "DP concentration hyperparameter (mass A)",
           y = "Metric",
            title = "Sensitivity to the DP concentration hyperparameter, dims = 5") +
      theme_bw()
  ggsave("bigSimDPhyperSens.png", width = 10, dpi = 300, units = "in")
} else {
  message("DP-hyperparameter sensitivity skipped: no dim2 CSVs with varying mass (mass A) found in the working directory.")
}
```

## Imbalanced-cluster scenarios
The simulation also includes scenarios with imbalanced cluster sizes (exponentially
decaying cluster frequencies). The table and figure below summarise model performance
on those scenarios using the same metrics as above: clustering (adjusted Rand index and
SALSO Binder/VI point estimates) and prediction (OOS RMSE and OOS log predictive score).

```{r}
#| column: page
# Model comparison table, imbalanced scenarios
imbalTbl <- longDF %>%
    filter(dims == 5, Imbalanced == TRUE) %>%
    group_by(Metric, Model) %>%
    summarise(Statistic = mean(Statistic, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = Model, values_from = Statistic)
imbalTbl |> write_csv("tables/table_model_comparison_imbalanced.csv")
imbalTbl |>
  gt() |>
  tab_header(title = "Imbalanced-cluster scenarios (dims=5)") |>
  fmt_number(decimals = 3, use_seps = FALSE) |>
  opt_interactive(use_compact_mode = TRUE, use_page_size_select = TRUE) |>
  gt_scroll()
```

```{r}
# Key-metric densities for imbalanced scenarios
longDF %>%
    filter(dims == 5, Imbalanced == TRUE, N == 1000, Common == 1.0,
           Metric %in% c("rind", "rmse", "binder", "lps")) %>%
    mutate(Metric = recode(Metric,
                           rind   = "Adjusted Rand Index",
                           rmse   = "OOS RMSE",
                           binder = "SALSO Binder ARI",
                           lps    = "OOS log predictive score")) %>%
    ggplot(aes(x = Statistic, fill = Model, y = after_stat(density))) +
    geom_histogram(alpha = 0.3, color = "black") +
    facet_grid(rows = vars(Metric), cols = vars(xdiff), scales = "free") +
    labs(x = "Statistic", y = "Density") +
    ggtitle("Imbalanced-cluster scenarios, N = 1000, dims = 5") +
    theme_bw()
ggsave("bigSimImbalanced.png", width = 8, dpi = 300, units = "in")
```

# common-effect CI coverage across all covariates
```{r}
#| column: page
commonCovTbl <- df %>%
    filter(N == 1000, dims == 5, Common == 1.0) %>%
    reframe(.by = c(xdiff, Imbalanced), `ppmx-common_ccov` = mean(`ppmx-common_ccov`, na.rm = TRUE)) %>%
    mutate(across(`ppmx-common_ccov`, round, 3))
commonCovTbl |> write_csv("tables/table_common_effect_cov.csv")
commonCovTbl |>
  gt() |>
  tab_header(title = "Common-effect joint coverage (ppmx-common, N=1000, dims=5)") |>
  fmt_number(decimals = 3, use_seps = FALSE) |>
  opt_interactive(use_compact_mode = TRUE, use_page_size_select = TRUE) |>
  gt_scroll()
```

