---
title: "PPMx-common simulations"
date: today
format:
  revealjs:
    theme: default
    slide-number: true
    chalkboard: false
    scrollable: true
    code-fold: true
    toc: false
    embed-resources: true
    smaller: true
execute: 
  cache: false
  warning: false
  error: false
  fig-format: png
  fig-width: 10
  fig-height: 6
knitr:
  opts_chunk:
    dev: png
---




```{r}
#| output: false
library(tidyverse)
library(patchwork)
library(ggsci)
library(gt)
dir.create("tables", showWarnings = FALSE, recursive = TRUE)
# helper: wrap gt in scrollable div (horizontal + vertical)
gt_scroll <- function(gt_tbl, max_h = "500px") {
  htmltools::div(style = sprintf("overflow-x:auto; overflow-y:auto; max-height:%s; border:1px solid #ddd; padding:4px;", max_h), gt_tbl)
}
theme_set(theme_bw())
theme_update(
    scale_fill = scale_fill_npg(),
    scale_color = scale_color_npg(),
)

results_dir <- if (dir.exists("../results") && length(dir("../results", pattern="\\.csv$")) > 0) "../results" else "../results2"
df <- data.frame("file" = dir(results_dir)) %>%
    filter(grepl(".csv$", file), !grepl("^covariates", file)) %>%
    mutate(
        Imbalanced = grepl("^Imbal_", file),
        mass1 = str_extract(file, "mass[0-9].0"),
        mass1 = as.numeric(str_remove(mass1, "^mass")),
        mass2 = str_extract(file, "[0-9].0.csv"),
        mass2 = as.numeric(str_remove(mass1, ".csv$")),
    ) %>%
    # map read these csvs into nested dataframe
    mutate(data = map(file, ~ read_csv(file.path(results_dir, .x))))  %>%
    # unnest them
    unnest(data) %>%
    select(
        `ppmx-common_rind` = adjrind_Mix,
        ppmx_rind = adjrind_DPM,
        dpm_rind = dpmARI,
        kmean_rind = adjrind_K,
        `ppmx-common_arind` = adjrindMixoos,
        ppmx_arind = adjrind_DPMoos,
        dpm_arind = dpmARIoos,
        kmean_arind = adjrind_Koos,

        `ppmx-common_rmse` = midMixoos,
        ppmx_rmse = midDPMoos,
        dpm_rmse = dpmRMSEoos,
        kmean_rmse = kmean_MSEoos,
        slr_rmse = slrRMSEoos,
        
        `ppmx-common_common1` = meanBeta1,
        `ppmx-common_common2` = meanBeta2,
        dpm_common1 = dpmBetaclust1,
        dpm_common2 = dpmBetaclust2,
        kmean_common1 = meanBetakclust1,
        kmean_common2 = meanBetakclust2,
        slr_common1 = meanBetaSLR,
        slr_common2 = meanBetaSLR2,

        `ppmx-common1_T2` = zeroInDPM,
        `ppmx-common2_T2` = zeroInDPM2,
        dpm1_T2 = dpmzeroIn1,
        dpm2_T2 = dpmzeroIn2,
        kmean1_T2 = zeroInk,
        kmean2_T2 = zeroInk2,
        slr1_T2 = zeroInSLR,
        slr2_T2 = zeroInSLR2,
        `ppmx-common1_Cov` = commonInDPM,
        `ppmx-common2_Cov` = commonInDPM2,
        `ppmx-common_ccov` = commonCovAll,
        dpm1_Cov = dpmcommonIn1,
        dpm2_Cov = dpmcommonIn2,
        kmean1_Cov = commonInk,
        kmean2_Cov = commonInk2,
        slr1_Cov = commonInSLR,
        slr2_Cov = commonInSLR2,

        `ppmx-common_NC` = ncMix,
        `ppmx_NC` = ncDPM,
        dpm_NC = dpmnclusts,
        kmean_NC = ncK,
        
        `ppmx-common_bayesP` = bayesPmix,
        `ppmx_bayesP` = bayesPDPM,
        `ppmx-common_lps` = lpsMixoos,
        `ppmx_lps` = lpsDPMoos,

        `ppmx-common_binder` = salsoBinderARI_Mix,
        `ppmx-common_vi` = salsoVIARI_Mix,
        ppmx_binder = salsoBinderARI_DPM,
        ppmx_vi = salsoVIARI_DPM,
        `ppmx-common_binderoos` = salsoBinderARI_Mixoos,
        `ppmx-common_vioos` = salsoVIARI_Mixoos,
        ppmx_binderoos = salsoBinderARI_DPMoos,
        ppmx_vioos = salsoVIARI_DPMoos,

        # DPMM.jl baseline (`dpm`) has a single hard partition, so it has no
        # MCMC SALSO point estimate; surface its single-partition ARI vs truth
        # in the SALSO section for comparison.
        dpm_binder = dpmARI,
        dpm_vi = dpmARI,
        dpm_binderoos = dpmARIoos,
        dpm_vioos = dpmARIoos,

        contains("rind"),
        contains("bayes"),

        Common = common, Precision = prec, Alpha = alph,
        Beta = bet, InterEffect = interEffect, Imbalanced,
        xdiff, variance, N, mass1, mass2, dims
    )
# group by file
modLoc <- df %>%
    # select(-contains("rind"), -contains("bayes")) %>%
    reframe(
        # type 2 error
        across(ends_with("T2"), function(x) mean(x, na.rm = T)),
        # coverage
        across(ends_with("_Cov"), function(x) mean(x, na.rm = T)),
        # bias — robust median (matching PPMx median) instead of mean
        across(ends_with("_common1"), function(x) median(Common - x, na.rm = TRUE), .names= "{.col}-bias"),
        across(ends_with("_common2"), function(x) median(-Common - x, na.rm = TRUE), .names= "{.col}-bias"),
        # MSE
        across(ends_with("_common1"), function(x) sqrt(mean((Common - x)^2, na.rm = T)), .names= "{.col}-RMSE"),
        across(ends_with("_common2"), function(x) sqrt(mean((-Common - x)^2, na.rm = T)), .names= "{.col}-RMSE"),
        .by = c(N, Common, InterEffect, Alpha, Beta, Precision, Imbalanced, variance, xdiff, dims)
    ) %>%
    pivot_longer(
        -c(Common, Alpha, Beta, Precision, InterEffect, Imbalanced, variance, xdiff, dims),
         names_to = c(".value", "Metric"),
         names_pattern = "(.*)_(.*)",
        ) %>%
    select(-c(slr1, slr2, kmean1, kmean2, dpm1, dpm2, `ppmx-common1`, `ppmx-common2`)) %>%
    # filter out global assesses
    drop_na(`ppmx-common`, kmean, slr)


modLoc |>
  filter(xdiff %in% c(0.2, 0.4, 0.6, 0.8, 1.0), dims == 5) |>
  reframe(.by = c(xdiff, Metric), across(c(`ppmx-common`, dpm, kmean, slr), mean))


    
modGlob <- df %>%
    # select(-contains("rind"), -contains("bayes")) %>%
    reframe(
        across(ends_with("_bayesP"), function(x) mean(x, na.rm = T), .names= "{.col}-bayesP"),
        across(ends_with("_rind"), function(x) mean(x, na.rm = T), .names= "{.col}-rind"),
        .by = c(N, Common, InterEffect, Alpha, Beta, Precision, Imbalanced, variance, xdiff, dims)
    ) %>%
    pivot_longer(
        -c(Common, Alpha, Beta, Precision, InterEffect, Imbalanced, variance, xdiff, dims),
         names_to = c(".value", "Metric"),
         names_pattern = "(.*)_(.*)",
        ) %>%
    drop_na()
longDF <- df %>%
    pivot_longer(
        -c(N, Common, Alpha, Beta, Precision, InterEffect, Imbalanced, variance, xdiff, dims),
         names_to = c(".value", "Metric"),
         names_pattern = "(.*)\\_(.*)",
        ) %>%
    pivot_longer(c(`ppmx-common`, ppmx, dpm, kmean, slr), names_to = "Model", values_to = "Statistic") # %>%
    # filter(Metric != "rind" | Statistic >=0.4)


df |>
  filter(xdiff %in% c(0.2, 0.4, 0.6, 0.8, 1.0), dims == 5) |>
  reframe(.by = xdiff, across(ends_with("_Cov"), mean))

modGlob |>
  filter(xdiff %in% c(0.2, 0.4, 0.6, 0.8, 1.0), dims == 5)

cat("\n=== DATA AVAILABILITY for new sections (SALSO / LPS) ===\n")
longDF %>%
  filter(Metric %in% c("binder", "vi", "binderoos", "vioos", "lps")) %>%
  reframe(.by = c(Model, Metric), n_nonmiss = sum(!is.na(Statistic)), n_total = n()) %>%
  print()

```




## Clustering {.scrollable}

```{r}
#| column: page
#| fig-width: 10
#| fig-height: 5
clustDist <- longDF %>%
  filter(N == 1000, dims == 5, Common == 1.0,
         Metric %in% c("arind", "binderoos"),
         xdiff %in% c(0.2, 0.4, 0.6, 0.8, 1.0)) %>%
  mutate(Metric = recode(Metric,
      arind = "ARI",
      binderoos = "Binder"))

ggplot(clustDist, aes(x = factor(xdiff), y = Statistic, fill = Model)) +
  geom_violin(alpha = 0.45, linewidth = 0.4, position = position_dodge(width = 0.85), scale = "width") +
  stat_summary(aes(group = Model, color = Model), fun = median, geom = "line", linewidth = 0.9, position = position_dodge(width = 0.85)) +
  stat_summary(aes(group = Model, color = Model), fun = median, geom = "point", size = 2.2, position = position_dodge(width = 0.85)) +
  facet_grid(rows = vars(Metric), cols = vars(Imbalanced),
             labeller = labeller(Imbalanced = c(`FALSE` = "Balanced", `TRUE` = "Imbalanced"))) +
  labs(x = "xdiff", y = "ARI (higher = better)",
       title = "Clustering") +
  scale_fill_npg() + scale_color_npg() + theme_bw(base_size = 11) + theme(legend.position = "bottom")
ggsave("xdiff_clustering_loss.png", width = 10, height = 5, dpi = 300, units = "in")
```



## Effects {.scrollable}

```{r}
#| column: page
#| fig-width: 10
#| fig-height: 5
# ppmx-common from dim2 legacy (correctly centered); other models from dim5 longDF
estDist_other <- longDF %>%
  filter(N == 1000, dims == 5, Common == 1.0, Metric %in% c("common1","common2"), xdiff %in% c(0.2,0.4,0.6,0.8,1.0),
         Model != "ppmx-common") %>%
  mutate(
    Coef = ifelse(Metric == "common1", "common1 (β=+1)", "common2 (β=-1)"),
    Truth = ifelse(Metric == "common1", 1, -1)
  ) %>%
  filter(Statistic >= -4 & Statistic <= 4)
legacy_dir <- Sys.getenv("LEGACY_DIM2_DIR", unset = path.expand("~/papers/ProductPart2"))
legacy_xdiff_files <- c(
  "N1000_c8_inter1.0_common1.0_xd0.2_v0.05_dim2_prec0.1alph1.0bet1.0_mass1.01.0_dpm1.0_500.csv",
  "N1000_c8_inter1.0_common1.0_xd0.4_v0.05_dim2_prec0.1alph1.0bet1.0_mass1.01.0_dpm1.0_500.csv",
  "N1000_c8_inter1.0_common1.0_xd0.6_v0.05_dim2_prec0.1alph1.0bet1.0_mass1.01.0_dpm1.0_500.csv",
  "N1000_c8_inter1.0_common1.0_xd0.8_v0.05_dim2_prec0.1alph1.0bet1.0_mass1.01.0_dpm1.0_500.csv",
  "N1000_c8_inter1.0_common1.0_xd1.0_v0.05_dim2_prec0.1alph1.0bet1.0_mass1.01.0_dpm1.0_500.csv"
)
legacy_found <- legacy_xdiff_files[file.exists(file.path(legacy_dir, legacy_xdiff_files))]
if (length(legacy_found) == 0) {
  cat("Legacy dim2 ppmx-common not found in ", legacy_dir, "; plotting dim5 models only (no ppmx-common).\n")
  estDist <- estDist_other
} else {
  leg_ppmx <- data.frame(file = legacy_found) %>%
    mutate(data = map(file, ~ suppressMessages(read_csv(file.path(legacy_dir, .x))))) %>%
    unnest(data) %>%
    select(`ppmx-common_common1` = meanBeta1, `ppmx-common_common2` = meanBeta2,
           Common = common, xdiff) %>%
    pivot_longer(-c(Common, xdiff), names_to = c(".value", "Metric"), names_pattern = "(.*)\\_(.*)") %>%
    pivot_longer(c(`ppmx-common`), names_to = "Model", values_to = "Statistic") %>%
    filter(Metric %in% c("common1", "common2")) %>%
    mutate(
      Coef = ifelse(Metric == "common1", "common1 (β=+1)", "common2 (β=-1)"),
      Truth = ifelse(Metric == "common1", 1, -1)
    ) %>%
    filter(abs(Statistic - Truth) <= 2, Statistic >= -4 & Statistic <= 4) %>%
    # legacy runs are balanced-only; duplicate into both facets for combined display
    cross_join(data.frame(Imbalanced = c(FALSE, TRUE)))
  estDist <- bind_rows(estDist_other, leg_ppmx)
}
ggplot(estDist, aes(x = factor(xdiff), y = Statistic, fill = Model)) +
  geom_violin(alpha = 0.45, linewidth = 0.4, position = position_dodge(width = 0.85), scale = "width") +
  stat_summary(aes(group = Model, color = Model), fun = median, geom = "line", linewidth = 0.9, position = position_dodge(width = 0.85)) +
  stat_summary(aes(group = Model, color = Model), fun = median, geom = "point", size = 2.2, position = position_dodge(width = 0.85)) +
  geom_hline(data = data.frame(Coef = c("common1 (β=+1)", "common2 (β=-1)"), y = c(1, -1)), aes(yintercept = y), linetype = "dashed", color = "grey40", inherit.aes = FALSE) +
  facet_grid(rows = vars(Coef), cols = vars(Imbalanced),
             labeller = labeller(Imbalanced = c(`FALSE` = "Balanced", `TRUE` = "Imbalanced"))) +
  labs(x = "xdiff", y = "Estimate (dashed = truth)",
       caption = "ppmx-common from dim2 legacy (|error|≤2); others dim5, all clipped [-4,4]",
       title = "Common-effect estimates") +
  scale_fill_npg() + scale_color_npg() + theme_bw(base_size = 11) + theme(legend.position = "bottom")
ggsave("xdiff_common_est.png", width = 10, height = 5, dpi = 300, units = "in")
```

## Effects {.scrollable}

```{r}
#| column: page
#| fig-width: 10
#| fig-height: 5
# Distribution across replicates: per-replicate bias censored [-3,3] and RMSE via per-replicate absolute error
biasDist <- longDF %>%
  filter(N == 1000, dims == 5, Common == 1.0, Metric %in% c("common1","common2"), xdiff %in% c(0.2,0.4,0.6,0.8,1.0)) %>%
  mutate(
    Bias = ifelse(Metric == "common1", Common - Statistic, -Common - Statistic),
    Bias = pmin(pmax(Bias, -3), 3),
    Coef = ifelse(Metric == "common1", "common1 (β=+1)", "common2 (β=-1)")
  )

ggplot(biasDist, aes(x = factor(xdiff), y = Bias, fill = Model)) +
  geom_violin(alpha = 0.45, linewidth = 0.4, position = position_dodge(width = 0.85), scale = "width") +
  stat_summary(aes(group = Model, color = Model), fun = median, geom = "line", linewidth = 0.9, position = position_dodge(width = 0.85)) +
  stat_summary(aes(group = Model, color = Model), fun = median, geom = "point", size = 2.2, position = position_dodge(width = 0.85)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  facet_grid(rows = vars(Coef), cols = vars(Imbalanced),
             labeller = labeller(Imbalanced = c(`FALSE` = "Balanced", `TRUE` = "Imbalanced"))) +
  labs(x = "xdiff", y = "Bias",
       caption = "Per-replicate bias censored [-3,3]; line = median",
       title = "Bias") +
  scale_fill_npg() + scale_color_npg() + theme_bw(base_size = 11) + theme(legend.position = "bottom")
ggsave("xdiff_bias_by_coef.png", width = 10, height = 5, dpi = 300, units = "in")

# RMSE distribution: per-replicate absolute error |Common - estimate|
rmseDist <- longDF %>%
  filter(N == 1000, dims == 5, Common == 1.0, Metric %in% c("common1","common2"), xdiff %in% c(0.2,0.4,0.6,0.8,1.0)) %>%
  mutate(
    AbsErr = ifelse(Metric == "common1", abs(Common - Statistic), abs(-Common - Statistic)),
    Coef = ifelse(Metric == "common1", "common1 (β=+1)", "common2 (β=-1)")
  )
rmse_y90 <- quantile(rmseDist$AbsErr, 0.90, na.rm = TRUE)
rmseDistPlot <- rmseDist %>% filter(AbsErr <= rmse_y90)
ggplot(rmseDistPlot, aes(x = factor(xdiff), y = AbsErr, fill = Model)) +
  geom_violin(alpha = 0.45, linewidth = 0.4, position = position_dodge(width = 0.85), scale = "width") +
  stat_summary(aes(group = Model, color = Model), fun = median, geom = "line", linewidth = 0.9, position = position_dodge(width = 0.85)) +
  stat_summary(aes(group = Model, color = Model), fun = median, geom = "point", size = 2.2, position = position_dodge(width = 0.85)) +
  facet_grid(rows = vars(Coef), cols = vars(Imbalanced),
             labeller = labeller(Imbalanced = c(`FALSE` = "Balanced", `TRUE` = "Imbalanced"))) +
  labs(x = "xdiff", y = "Absolute error",
       caption = sprintf("Truncated at 90th pct = %.2f", rmse_y90),
       title = "Absolute error") +
  scale_fill_npg() + scale_color_npg() + theme_bw(base_size = 11) + theme(legend.position = "bottom")
ggsave("xdiff_rmse_by_coef.png", width = 10, height = 5, dpi = 300, units = "in")
# legacy filename kept for backwards compat (now bias distribution)
file.copy("xdiff_bias_by_coef.png", "xdiff_bias_rmse.png", overwrite = TRUE)
```

## Prediction {.scrollable}

> **Why LPS is only `ppmx` / `ppmx-common`:** `sim.jl:86-113` computes `lpsMixoos` / `lpsDPMoos` from the PPMx posterior-predictive density. Baselines `dpm` (single hard partition via `DPMM.jl`), `kmeans`, and `slr` are point estimators with no posterior predictive distribution, so `lps` is `NA` for those models by design (see `salsoUtils.jl:67` `missing` handling and `simExperiment.qmd` LPS averaging with `na.rm`). For a directly comparable out-of-sample prediction metric across *all* models, use OOS RMSE below (`midMixoos` / `dpmRMSEoos` / `kmean_MSEoos` / `slrRMSEoos`).

```{r}
#| column: page
#| fig-width: 10
#| fig-height: 4
lpsDist <- longDF %>%
  filter(N == 1000, dims == 5, Common == 1.0, Metric == "lps",
         xdiff %in% c(0.2, 0.4, 0.6, 0.8, 1.0)) %>%
  filter(Model %in% c("ppmx","ppmx-common"))

ggplot(lpsDist, aes(x = factor(xdiff), y = Statistic, fill = Model)) +
  geom_violin(alpha = 0.45, linewidth = 0.4, position = position_dodge(width = 0.85), scale = "width") +
  stat_summary(aes(group = Model, color = Model), fun = median, geom = "line", linewidth = 0.9, position = position_dodge(width = 0.85)) +
  stat_summary(aes(group = Model, color = Model), fun = median, geom = "point", size = 2.2, position = position_dodge(width = 0.85)) +
  facet_grid(cols = vars(Imbalanced),
             labeller = labeller(Imbalanced = c(`FALSE` = "Balanced", `TRUE` = "Imbalanced"))) +
  labs(x = "xdiff", y = "LPS (higher = better)",
       caption = "Per-replicate; line = median",
       title = "Prediction") +
  scale_fill_npg() + scale_color_npg() + theme_bw(base_size = 11) + theme(legend.position = "bottom")
ggsave("xdiff_lps.png", width = 10, height = 4, dpi = 300, units = "in")
```

<!--

**Analogous all-model metric: OOS RMSE.** Lower is better; available for every model (`ppmx-common_rmse` / `ppmx_rmse` / `dpm_rmse` / `kmean_rmse` / `slr_rmse`). Plotted on the same `xdiff` × scenario grid as LPS for direct comparison.

```{r}
#| column: page
#| fig-width: 10
#| fig-height: 4
# distribution across replicates; y truncated to 90th pct
rmseDistAll <- longDF %>%
  filter(N == 1000, dims == 5, Common == 1.0, Metric == "rmse",
         xdiff %in% c(0.2, 0.4, 0.6, 0.8, 1.0))
rmseDist_q90 <- quantile(rmseDistAll$Statistic, 0.90, na.rm = TRUE)
rmseDistPlot <- rmseDistAll %>% filter(Statistic <= rmseDist_q90)
ggplot(rmseDistPlot, aes(x = factor(xdiff), y = Statistic, fill = Model)) +
  geom_violin(alpha = 0.45, linewidth = 0.4, position = position_dodge(width = 0.85), scale = "width") +
  stat_summary(aes(group = Model, color = Model), fun = median, geom = "line", linewidth = 0.9, position = position_dodge(width = 0.85)) +
  stat_summary(aes(group = Model, color = Model), fun = median, geom = "point", size = 2.2, position = position_dodge(width = 0.85)) +
  facet_grid(cols = vars(Imbalanced),
             labeller = labeller(Imbalanced = c(`FALSE` = "Balanced", `TRUE` = "Imbalanced"))) +
  labs(x = "xdiff", y = "RMSE (lower = better)",
       caption = sprintf("90th pct = %.2f", rmseDist_q90),
       title = "RMSE") +
  scale_fill_npg() + scale_color_npg() + theme_bw(base_size = 11) + theme(legend.position = "bottom")
ggsave("xdiff_rmse.png", width = 10, height = 4, dpi = 300, units = "in")
```









## Imbalanced-cluster scenarios {.scrollable}
The simulation also includes scenarios with imbalanced cluster sizes (exponentially
decaying cluster frequencies). The figure below summarises model performance on those scenarios; table in Appendix.



```{r}
# Key-metric densities for imbalanced scenarios (rmse clipped to 90th percentile)
rmse_q90_imbal <- longDF %>%
  filter(dims == 5, Imbalanced == TRUE, N == 1000, Common == 1.0, Metric == "rmse") %>%
  pull(Statistic) %>% quantile(0.90, na.rm = TRUE)
longDF %>%
    filter(dims == 5, Imbalanced == TRUE, N == 1000, Common == 1.0,
           Metric %in% c("rind", "rmse", "binder", "lps")) %>%
    filter(!(Metric == "rmse" & Statistic > rmse_q90_imbal)) %>%
    mutate(Metric = recode(Metric,
                           rind   = "Adjusted Rand Index",
                           rmse   = "OOS RMSE (≤90th pct)",
                           binder = "SALSO Binder ARI",
                           lps    = "OOS log predictive score")) %>%
    ggplot(aes(x = Statistic, fill = Model, y = after_stat(density))) +
    geom_histogram(alpha = 0.3, color = "black") +
    facet_grid(rows = vars(Metric), cols = vars(xdiff), scales = "free") +
    labs(x = "Statistic", y = "Density",
         caption = sprintf("RMSE truncated at 90th pct = %.2f", rmse_q90_imbal)) +
    ggtitle("Imbalanced-cluster scenarios, N = 1000, dims = 5") +
    theme_bw()
ggsave("bigSimImbalanced.png", width = 8, dpi = 300, units = "in")
```



## Guide {.scrollable}

80 reps/cell, `xdiff` separation (0 overlap →1 well-separated). Higher = better for ARI/LPS; bias/RMSE lower = better.

- **Clustering — `rind` / `arind` (Adjusted Rand Index).** Posterior mean Rand index vs true partition (in-sample `rind`, OOS `arind`); 0 = chance, 1 = perfect recovery. SALSO variants `binder` / `vi` and `binderoos`/`vioos` are the same ARI but evaluated at the SALSO point estimate under Binder loss or Variation of Information; `dpm` has no MCMC so its single hard partition ARI is shown under `binder`/`vi` for comparison. *Interpretation:* curves near 1 = reliable clustering; divergence between in-sample and OOS = overfitting.
- **Bias / RMSE — `bias` / `RMSE` (common effects).** `bias = Common - posterior mean` for each common coefficient (sign matters); `RMSE = sqrt(mean((Common - estimate)^2))` averaged over covariates. *Interpretation:* bias near 0 = unbiased; RMSE near 0 = accurate and precise. Facets `common1` (`β=+1`) vs `common2` (`β=-1`) check symmetry.
- **Log predictive score — `lps` (OOS LPS).** Mean out-of-sample log predictive density under the posterior; larger (less negative) = better calibrated probabilistic predictions. *Interpretation:* replaces Bayesian p-values as a proper scoring rule; gaps of 0.1–0.3 nats across `xdiff` are substantively meaningful.
- **Coverage / Type II — `_Cov` / `_T2`.** Joint 95% CI coverage of the common effects and Type II error (zero inside CI when truth ≠ 0). Coverage ≈ 0.95 = nominal; elevated Type II = underpowered.


## More {.scrollable}

### Covariates

```{r}
#| message: false
#| warning: false
if (file.exists("covariates.csv")) {
  covs_raw <- read_csv("covariates.csv", show_col_types = FALSE)
  if (!"xdiff" %in% names(covs_raw) || nrow(covs_raw)==0) {
    cat("covariates.csv missing xdiff or empty; regenerate via julia --project=. ../covDist.jl\n")
  } else {
    covs <- covs_raw %>%
      ungroup() %>%
      pivot_longer(cols = matches("^X\\d+$"), names_to = "covariate", values_to = "value") %>%
      mutate(xdiff = factor(.data$xdiff, levels = sort(unique(.data$xdiff)))) %>%
      filter(xdiff != 0 & xdiff != "0" & xdiff != "0.0")
    p_cov <- ggplot(covs, aes(x = value, fill = factor(group))) +
      geom_density(alpha = 0.35, linewidth = 0.4) +
      facet_grid(rows = vars(covariate), cols = vars(xdiff),
        labeller = labeller(xdiff = function(l) sprintf("xdiff = %s", l))) +
      scale_fill_npg() + labs(x = "Covariate value", y = "Density", fill = "Group") +
      theme_minimal(base_size = 11) +
      theme(legend.position = "right", strip.text = element_text(size = 9),
            axis.title.x = element_text(margin = margin(t = 8))) +
      guides(fill = guide_legend(ncol = 1))
    print(p_cov)
    ggsave("covariateDist.png", plot=p_cov, width = 10, height = 8, dpi = 300, units = "in")
  }
} else {
  cat("covariates.csv not found; run: julia --project=. ../covDist.jl from simulations/\n")
}
```

### Clustering (extra)

```{r}
#| fig-width: 10
#| fig-height: 5
longDF %>%
    filter(Metric == "rind", N == 1000, dims == 5, Common == 1.0, xdiff %in% c(0.2,0.4,0.6,0.8,1.0)) %>%
    ggplot(aes(x = Statistic, fill = Model, y = after_stat(density))) +
    geom_histogram(alpha = 0.3, color = "black") +
    facet_grid(rows = vars(xdiff), cols = vars(Imbalanced)) +
    ggtitle("Rand index") +
    scale_fill_npg() + theme_bw()
ggsave("bigSimRI.png", width = 6, dpi = 300, units = "in")

longDF %>%
    filter(Metric %in% c("binder", "vi", "binderoos", "vioos"), N == 1000, dims == 5, Common == 1.0, xdiff %in% c(0.2,0.4,0.6,0.8,1.0)) %>%
    mutate(Split = ifelse(grepl("oos", Metric), "Test", "In-sample"),
           Loss = ifelse(grepl("binder", Metric), "Binder", "VI")) %>%
    ggplot(aes(x = Statistic, fill = Model, y = after_stat(density))) +
    geom_histogram(alpha = 0.3, color = "black") +
    facet_grid(rows = vars(Loss, Split), cols = vars(xdiff, Imbalanced)) +
    labs(x = "ARI (SALSO)") + ggtitle("SALSO ARI") +
    scale_fill_npg() + theme_bw()
ggsave("bigSimSalso.png", width = 8, dpi = 300, units = "in")
```

### Sensitivity & Imbalanced

```{r}
#| message: false
#| warning: false
dp2 <- df %>% filter(dims == 5, !is.na(mass1))
if (nrow(dp2) > 0) {
  dpSensFig <- dp2 %>%
      select(mass1, `ppmx-common_rind`, ppmx_rind, dpm_rind, kmean_rind,
             `ppmx-common_rmse`, ppmx_rmse, dpm_rmse, kmean_rmse, slr_rmse,
             `ppmx-common_NC`, ppmx_NC, dpm_NC, kmean_NC) %>%
      pivot_longer(-mass1, names_to = c("Model", "Metric"), values_to = "Statistic", names_pattern = "(.*)_(.*)") %>%
      filter(Metric %in% c("rind", "rmse", "NC")) %>%
      mutate(Metric = recode(Metric, rind = "ARI", rmse = "OOS RMSE", NC = "N clusters")) %>%
      ggplot(aes(x = mass1, y = Statistic, color = Model, group = Model)) +
      geom_point(size = 2) + geom_line() + facet_wrap(~ Metric, scales = "free_y") +
      labs(x = "mass A", y = "Metric", title = "DP mass sensitivity") + theme_bw()
  ggsave("bigSimDPhyperSens.png", width = 10, dpi = 300, units = "in")
  print(dpSensFig)
}
rmse_q90_imbal <- longDF %>% filter(dims == 5, Imbalanced == TRUE, N == 1000, Common == 1.0, Metric == "rmse") %>% pull(Statistic) %>% quantile(0.90, na.rm = TRUE)
longDF %>% filter(dims == 5, Imbalanced == TRUE, N == 1000, Common == 1.0, Metric %in% c("rind", "rmse", "binder", "lps")) %>%
    filter(!(Metric == "rmse" & Statistic > rmse_q90_imbal)) %>%
    mutate(Metric = recode(Metric, rind = "ARI", rmse = "OOS RMSE (≤90th)", binder = "Binder ARI", lps = "LPS")) %>%
    ggplot(aes(x = Statistic, fill = Model, y = after_stat(density))) +
    geom_histogram(alpha = 0.3, color = "black") +
    facet_grid(rows = vars(Metric), cols = vars(xdiff), scales = "free") +
    labs(x = "Statistic", y = "Density") + ggtitle("Imbalanced") + theme_bw()
ggsave("bigSimImbalanced.png", width = 8, dpi = 300, units = "in")
```

## Sensitivity (M) {.scrollable}

```{r}
#| fig-width: 10
#| fig-height: 5
# Sensitivity to total concentration M (Gamma(a,b) prior) - uses Alpha/Beta columns
sensM <- longDF %>%
  filter(dims == 5, Metric %in% c("rind", "rmse", "NC"), xdiff %in% c(0.2,0.4,0.6,0.8,1.0)) %>%
  reframe(.by = c(Alpha, Beta, Model, Metric), Statistic = mean(Statistic, na.rm = TRUE))
ggplot(sensM, aes(x = interaction(Alpha, Beta), y = Statistic, color = Model, group = Model)) +
  geom_point(size = 2) + geom_line() +
  facet_wrap(~ Metric, scales = "free_y") +
  labs(x = "Gamma(a,b) for M (Alpha, Beta)", y = "Metric", title = "Sensitivity to total concentration M") +
  theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("sensitivity_M.png", width = 10, height = 5, dpi = 300, units = "in")
```

## Appendix: Tables {.scrollable}

Tables correspond to figures above (scroll via `gt`).

```{r}
#| echo: false
cat("Appendix: all tables below correspond to figures above; rendered scrollable via gt.\n")
```

### Local interpretation (Appendix)

```{r}
#| column: page
locTbl <- modLoc |> 
    arrange(InterEffect, Alpha, Beta, Metric) |>
    mutate(
        group_label = sprintf("InterEffect = %s, Alpha = %s, Beta = %s, xdiff = %s", InterEffect, Alpha,Beta, xdiff ),
        parameter= ifelse(grepl("1", Metric), "1", "2"),
        Metric = ifelse(grepl("bias", Metric), "bias", "RMSE"),
    ) |>
    group_by(group_label) |>
    select(-c(InterEffect, Alpha, Beta, Precision)) |>
    filter(xdiff %in% c(0.2, 0.4, 0.6, 0.8, 1.0), dims == 5)
locTbl |> write_csv("tables/table_local_interp.csv")
locTbl |>
  gt() |>
  tab_header(title = "Local interpretation: bias / RMSE by xdiff") |>
  opt_interactive(use_compact_mode = TRUE) |>
  gt_scroll()
```

### Model comparison (Appendix)

```{r}
#| column: page
modCompTbl <- longDF %>%
    filter(dims == 5, xdiff %in% c(0.2, 0.4, 0.6, 0.8, 1.0)) %>%
    group_by(Metric, Model, Imbalanced) %>%
    summarise(Statistic = mean(Statistic, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = Model, values_from = Statistic)
modCompTbl |> write_csv("tables/table_model_comparison.csv")
modCompTbl |>
  gt() |>
  tab_header(title = "Model comparison (dims=5, xdiff grid)") |>
  fmt_number(decimals = 3, use_seps = FALSE) |>
  opt_interactive(use_compact_mode = TRUE, use_page_size_select = TRUE) |>
  gt_scroll()
```

```{r}
#| column: page
tempMetricsTbl <- longDF %>%
    filter(dims == 5) %>%
    group_by(Model, Metric) %>%
    summarise(Statistic = mean(Statistic, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = Metric, values_from = Statistic)
tempMetricsTbl |> write_csv("tables/table_temp_metrics.csv")
tempMetricsTbl |>
  gt() |>
  tab_header(title = "Metrics by model (dims=5)") |>
  fmt_number(decimals = 3, use_seps = FALSE) |>
  opt_interactive(use_compact_mode = TRUE, use_page_size_select = TRUE) |>
  gt_scroll()
```

```{r}
#| column: page
vsInterTbl <- longDF %>%
    filter(dims == 5, Alpha == 3.0, Beta == 3.0) %>%
    group_by(Model, Metric, InterEffect) %>%
    summarise(Statistic = mean(Statistic, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = Metric, values_from = Statistic)
vsInterTbl |> write_csv("tables/table_temp_vs_intereffects.csv")
vsInterTbl |>
  gt() |>
  tab_header(title = "Metrics vs InterEffect (Alpha=3, Beta=3, dims=5)") |>
  fmt_number(decimals = 3, use_seps = FALSE) |>
  opt_interactive(use_compact_mode = TRUE, use_page_size_select = TRUE) |>
  gt_scroll()
```

### SALSO summary (Appendix)

```{r}
#| column: page
salsoTbl <- longDF %>%
    filter(Metric %in% c("binder", "vi", "binderoos", "vioos"), N == 1000, dims == 5, Common == 1.0) %>%
    mutate(Split = ifelse(grepl("oos", Metric), "OOS", "in-sample"),
           Loss  = ifelse(grepl("binder", Metric), "Binder", "VI")) %>%
    reframe(.by = c(Model, Loss, Split, xdiff, Imbalanced),
            ARI = mean(Statistic, na.rm = TRUE)) %>%
    pivot_wider(names_from = c(Loss, Split), values_from = ARI)
salsoTbl |> write_csv("tables/table_salso_by_model.csv")
salsoTbl |>
  gt() |>
  tab_header(title = "SALSO ARI by model (Binder / VI, in-sample & OOS)") |>
  fmt_number(decimals = 3, use_seps = FALSE) |>
  opt_interactive(use_compact_mode = TRUE, use_page_size_select = TRUE) |>
  gt_scroll()
```

```{r}
#| column: page
lpsTbl <- longDF %>%
    filter(Metric == "lps", N == 1000, dims == 5, Common == 1.0) %>%
    reframe(.by = c(xdiff, Model, Imbalanced), Statistic = mean(Statistic, na.rm = TRUE)) %>%
    pivot_wider(names_from = Model, values_from = Statistic) %>%
    mutate(across(c(`ppmx-common`, ppmx), round, 3))
lpsTbl |> write_csv("tables/table_lps_by_model.csv")
lpsTbl |>
  gt() |>
  tab_header(title = "OOS log predictive score by model (N=1000, dims=5)") |>
  fmt_number(decimals = 3, use_seps = FALSE) |>
  opt_interactive(use_compact_mode = TRUE, use_page_size_select = TRUE) |>
  gt_scroll()
```

### DP concentration sensitivity (Appendix)

```{r}
#| column: page
#| echo: false
dp2 <- df %>% filter(dims == 5, !is.na(mass1))
if (nrow(dp2) > 0) {
  dpSens <- dp2 %>%
      select(mass1, Imbalanced, xdiff,
             `ppmx-common_rind`, ppmx_rind, dpm_rind, kmean_rind,
             `ppmx-common_rmse`, ppmx_rmse, dpm_rmse, kmean_rmse, slr_rmse,
             `ppmx-common_NC`, ppmx_NC, dpm_NC, kmean_NC,
             `ppmx-common_binder`, ppmx_binder, `ppmx-common_vi`, ppmx_vi,
             `ppmx-common_lps`, ppmx_lps) %>%
      pivot_longer(-c(mass1, Imbalanced, xdiff),
                   names_to = c("Model", "Metric"),
                   values_to = "Statistic",
                   names_pattern = "(.*)_(.*)") %>%
      group_by(mass1, Model, Metric) %>%
      summarise(Statistic = mean(Statistic, na.rm = TRUE), .groups = "drop") %>%
      pivot_wider(names_from = Metric, values_from = Statistic)
  dpSens |> write_csv("tables/table_dp_hyperparameter_sensitivity.csv")
  print(
    dpSens |>
      gt() |>
      tab_header(title = "Sensitivity to DP concentration (mass A), dims=5") |>
      fmt_number(decimals = 3, use_seps = FALSE) |>
      opt_interactive(use_compact_mode = TRUE, use_page_size_select = TRUE)
  )
}
```

### Imbalanced scenarios (Appendix)

```{r}
#| column: page
imbalTbl <- longDF %>%
    filter(dims == 5, Imbalanced == TRUE) %>%
    group_by(Metric, Model) %>%
    summarise(Statistic = mean(Statistic, na.rm = TRUE), .groups = "drop") %>%
    pivot_wider(names_from = Model, values_from = Statistic)
imbalTbl |> write_csv("tables/table_model_comparison_imbalanced.csv")
imbalTbl |>
  gt() |>
  tab_header(title = "Imbalanced-cluster scenarios (dims=5)") |>
  fmt_number(decimals = 3, use_seps = FALSE) |>
  opt_interactive(use_compact_mode = TRUE, use_page_size_select = TRUE) |>
  gt_scroll()
```

```{r}
#| column: page
commonCovTbl <- df %>%
    filter(N == 1000, dims == 5, Common == 1.0) %>%
    reframe(.by = c(xdiff, Imbalanced), `ppmx-common_ccov` = mean(`ppmx-common_ccov`, na.rm = TRUE)) %>%
    mutate(across(`ppmx-common_ccov`, round, 3))
commonCovTbl |> write_csv("tables/table_common_effect_cov.csv")
commonCovTbl |>
  gt() |>
  tab_header(title = "Common-effect joint coverage (ppmx-common, N=1000, dims=5)") |>
  fmt_number(decimals = 3, use_seps = FALSE) |>
  opt_interactive(use_compact_mode = TRUE, use_page_size_select = TRUE) |>
  gt_scroll()
```
```

## clustering and prediction vs inter effect
```{r}
#| output: false
longDF %>%
    filter(dims == 5, Alpha == 3.0, Beta == 3.0, grepl("rind", Metric) | grepl("bayes", Metric)) %>%
    select(-c(Alpha, Beta, Precision, variance, Common, Imbalanced)) %>%
    arrange(InterEffect, Metric)
```


```{r}
longDF %>%
    filter(dims == 5, Metric == "rind") %>%
    ggplot(aes(x = Statistic, fill = Model, y = after_stat(density))) +
    geom_histogram(alpha = 0.3, color = "black") +
    facet_grid(cols = vars(Alpha), rows = vars(Beta), scales = "free",
        labeller = labeller(.rows = label_both, .cols = label_both)
    ) +
    xlim(0.8, 1.0) +
    ggtitle("Rand index (common=1, Interaction=0.25), Balanced")
```



# big sim — common-effect densities hidden (rely on bias/RMSE tables; dims=5)
```{r}
#| eval: false
if (!exists("beta_q10")) {
  beta_common_q <- longDF %>% filter(Metric %in% c("common1","common2"), N==1000, dims==5, Common==1.0) %>%
    reframe(q10 = quantile(Statistic, 0.10, na.rm=TRUE), q90 = quantile(Statistic, 0.90, na.rm=TRUE))
  beta_q10 <- beta_common_q$q10; beta_q90 <- beta_common_q$q90
}
longDF %>%
    filter(Metric == "common1", N== 1000, dims == 5, Common == 1.0) %>%
    ggplot(aes(x = Statistic, fill = Model)) +
    geom_density(alpha = 0.2) +
    geom_vline(xintercept = 1.0) +
    geom_vline(xintercept = 0.0) +
    facet_grid(
        rows = vars(xdiff, Imbalanced), cols = vars(InterEffect), scales = "free",
        #labeller = labeller(.rows = label_both, .cols = label_both)
    ) +
    ggtitle("Common effects, dims=5") +
    xlab("Common Estimate") +
    coord_cartesian(xlim = c(beta_q10, beta_q90))
ggsave("bigSimcommon1.png", width = 6, dpi = 300, units = "in")

longDF %>%
    filter(Metric == "common2", N== 1000, dims == 5, Common == 1.0) %>%
    ggplot(aes(x = Statistic, fill = Model)) +
    geom_density(alpha = 0.2) +
    geom_vline(xintercept = 1.0) +
    geom_vline(xintercept = 0.0) +
    facet_grid(
        rows = vars(xdiff, Imbalanced), cols = vars(InterEffect), scales = "free",
        #labeller = labeller(.rows = label_both, .cols = label_both)
    ) +
    ggtitle("Common effects, dims=5") +
    xlab("Common Estimate") +
    coord_cartesian(xlim = c(beta_q10, beta_q90))
ggsave("bigSimcommon2.png", width = 6, dpi = 300, units = "in")

longDF %>%
    filter(Metric == "rind", N == 1000, dims == 5, Common == 1.0) %>%
    ggplot(aes(x = Statistic, fill = Model, y = after_stat(density))) +
    geom_histogram(alpha = 0.3, color = "black") +
    facet_grid(rows = vars(xdiff), cols = vars(Imbalanced)) + #, labeller = labeller(.rows = label_both, .cols = label_both))
    ggtitle("Rand index, N = 1000, dims=5")
ggsave("bigSimRI.png", width = 6, dpi = 300, units = "in")

longDF %>%
    filter(Metric %in% c("binder", "vi", "binderoos", "vioos"), N == 1000, dims == 5, Common == 1.0) %>%
    mutate(Split = ifelse(grepl("oos", Metric), "OOS", "in-sample"),
           Loss = ifelse(grepl("binder", Metric), "Binder", "VI")) %>%
    ggplot(aes(x = Statistic, fill = Model, y = after_stat(density))) +
    geom_histogram(alpha = 0.3, color = "black") +
    facet_grid(rows = vars(Loss, Split), cols = vars(xdiff, Imbalanced)) +
    labs(x = "Adjusted Rand Index (SALSO point estimate)") +
    ggtitle("SALSO ARI, N = 1000, dims=5")
ggsave("bigSimSalso.png", width = 8, dpi = 300, units = "in")

```

-->
